# Results

## Fetal Deaths By Time
```{r}
library(ggplot2)
library(dplyr)
library(plotly)

### Total
g1 <- fetal %>%
  group_by(Year) %>%
  summarise(add = sum(Fetal.Deaths), add_p = sum(Percent.of.Total.Deaths)) %>%
  ggplot(aes(x = Year, y = add_p)) +
  geom_line() +
  geom_point(alpha = 0.5) +
  geom_point(aes(x = 2007, y = 7.22), color = "red") +
  geom_text(aes(label = add_p), nudge_x = 0.4, nudge_y = -0.06, size = 3) +
  labs(x = "Year", y = "Total Fetal Death Rate",
       title = "Fetal Death Rate Over Years") +
  scale_x_continuous(breaks = seq(2005, 2019, 1))
g1


g2 <- fetal %>%
  group_by(Sex, Year) %>%
  summarise(add_g = sum(Fetal.Deaths), add_g_p = sum(Percent.of.Total.Deaths)) %>%
  ggplot(aes(x = Year, y = add_g_p, col = Sex)) +
  geom_line() +
  geom_point(alpha = 0.5) +
  labs(x = "Year", y = "Fetal Death Rate",
       title = "Fetal Death Rate by Different Genders Over Years") +
  scale_x_continuous(breaks = seq(2005, 2019, 1))

start = 2007
end = 2008
g2 + annotate("rect", xmin = start, xmax = end, ymin = -Inf, ymax = Inf, 
              fill = "green", alpha = .2) +
  annotate("text", x = 2008.1, y = 3, label = "2007-2008", 
           color = "black", hjust = 0) +
  theme_classic()

# g3 <- birth_s_g %>%
#   group_by(Gender, Year) %>%
#   summarise(sum = sum(Births)) %>%
#   ggplot(aes(x = Year, y = sum, col = Gender)) +
#   geom_line() +
#   geom_point() +
#   labs(x = "Year", y = "Number of Births", 
#        title = "Number of Births by Different Genders Over Years") +
#   theme_classic() +
#   scale_x_continuous(breaks = seq(2007, 2019, 1))
# g3

# fetal %>%
#   group_by(Year) %>%
#   summarise(add = sum(Fetal.Deaths), add_p = sum(Percent.of.Total.Deaths)) %>%
#   plot_ly(x = ~Year, y = ~add, type = "scatter", mode = "lines+markers",
#           hoverinfo = "text", text = ~add) %>%
#   layout(title = "Total Fetal Deaths Over Years", 
#          xaxis = list(title = "Year", breaks = seq(2005, 2019, 1)),
#          yaxis = list(title = "Number of Total Death"))
```

First of all, let us take a look at how the fetal death rates changed between 2005 and 2019. We can clearly see a decreasing trend since 2007. Since we don’t have data before 2007, even though our graphs show there might be an increasing trend before 2007, we cannot simply make a conclusion. After further checking the movement by splitting the data into different genders, we found out that from 2007 to 2008, the female fetal death rate had a sharper decrease than male data did. This decreasing trend is a good indicator, showing an improvement in national health. It is also unsurprising: with the development of technology and medical skills, we can now detect any potential factors that lead to fetus deaths more efficiently.


## Distributions of Fetal Deaths

## Remarkable Fetal Characteristics

### Most Relevant

### Flow of Change

## Maternal/Paternal Characteristics

```{r}
library(tidyverse)
library(ggplot2)
parentsAge <- read.delim("parentsAge.txt") %>% 
  dplyr::select(Age.of.Mother.9, Age.of.Father,Fetal.Deaths) %>%
  filter(Age.of.Father != "" & Age.of.Father != "Unknown or Not Stated" & Age.of.Father != "Not Reported")
g_age <- ggplot(parentsAge, aes(x = Age.of.Father, y = Fetal.Deaths)) +
  geom_col(position = "dodge") +
  coord_flip() + 
  facet_wrap(~Age.of.Mother.9) + 
  ggtitle("Fetal Deaths Faceted by Maternal Age") + 
  xlab("Paternal Age") +
  ylab("# of Fetal Deaths")
print(g_age)
```

```{r}
race <- fetal %>% dplyr::select(Mother.s.Bridged.Race, Year, Fetal.Deaths) %>%
  filter(Mother.s.Bridged.Race != "Not Reported") %>%
  group_by(Mother.s.Bridged.Race, Year) %>%
  summarize(n = n())

library(ggridges)
g_race <- ggplot(race, aes(x = Year, y = Mother.s.Bridged.Race, fill = Mother.s.Bridged.Race)) +
  geom_density_ridges(aes(n)) +
  theme_ridges() + 
  theme(legend.position = "none") +
  ggtitle("")
print(g_race)
```



```{r}
Race_N_Marriage <- read.delim("RaceNMarriage.txt") %>% 
  rename('Maternal Race' = Mother.s.Bridged.Race) %>%
  rename('Maternal Marital Status' = Mother.s.Marital.Status) %>% 
  na.omit() %>%
  dplyr::select(`Maternal Race`, `Maternal Marital Status`, Fetal.Deaths) %>%
  filter(`Maternal Marital Status` == "Married" | `Maternal Marital Status` == "Unmarried")



```


### Clevland plot (age vs fetal death, facet by marital status)



