# Results

## Fetal Deaths Over Years
```{r}
library(ggplot2)
library(dplyr)
library(plotly)

### Total
g1 <- fetal %>%
  dplyr::group_by(Year) %>%
  dplyr::summarise(add = sum(Fetal.Deaths), add_p = sum(Percent.of.Total.Deaths)) %>%
  ggplot(aes(x = Year, y = add_p)) +
  geom_line() +
  geom_point(alpha = 0.5) +
  geom_point(aes(x = 2007, y = 7.22), color = "red") +
  geom_text(aes(label = add_p), nudge_x = 0.4, nudge_y = -0.06, size = 3) +
  labs(x = "Year", y = "Total Fetal Death Rate",
       title = "Fetal Death Rate Over Years") +
  scale_x_continuous(breaks = seq(2005, 2019, 1))
g1


g2 <- fetal %>%
  dplyr::group_by(Sex, Year) %>%
  dplyr::summarise(add_g = sum(Fetal.Deaths), add_g_p = sum(Percent.of.Total.Deaths)) %>%
  ggplot(aes(x = Year, y = add_g_p, col = Sex)) +
  geom_line() +
  geom_point(alpha = 0.5) +
  labs(x = "Year", y = "Fetal Death Rate",
       title = "Fetal Death Rate by Different Genders Over Years") +
  scale_x_continuous(breaks = seq(2005, 2019, 1))

start = 2007
end = 2008
g2 + annotate("rect", xmin = start, xmax = end, ymin = -Inf, ymax = Inf, 
              fill = "green", alpha = .2) +
  annotate("text", x = 2008.1, y = 3, label = "2007-2008", 
           color = "black", hjust = 0) +
  theme_classic()

# g3 <- birth_s_g %>%
#   group_by(Gender, Year) %>%
#   summarise(sum = sum(Births)) %>%
#   ggplot(aes(x = Year, y = sum, col = Gender)) +
#   geom_line() +
#   geom_point() +
#   labs(x = "Year", y = "Number of Births", 
#        title = "Number of Births by Different Genders Over Years") +
#   theme_classic() +
#   scale_x_continuous(breaks = seq(2007, 2019, 1))
# g3

# fetal %>%
#   group_by(Year) %>%
#   summarise(add = sum(Fetal.Deaths), add_p = sum(Percent.of.Total.Deaths)) %>%
#   plot_ly(x = ~Year, y = ~add, type = "scatter", mode = "lines+markers",
#           hoverinfo = "text", text = ~add) %>%
#   layout(title = "Total Fetal Deaths Over Years", 
#          xaxis = list(title = "Year", breaks = seq(2005, 2019, 1)),
#          yaxis = list(title = "Number of Total Death"))
```

First of all, let us take a look at how the fetal death rates changed between 2005 and 2019. We can clearly see a decreasing trend since 2007. Since we don’t have data before 2007, even though our graphs show there might be an increasing trend before 2007, we cannot simply make a conclusion. After further checking the movement by splitting the data into different genders, we found out that from 2007 to 2008, the female fetal death rate had a sharper decrease than male data did. We should notice that our graph shows higher death rates for male fetuses. This is not because male fetuses have a higher possibility of death but because male fetuses have higher birth rates in general. This decreasing trend is a good indicator, showing an improvement in national health. It is also unsurprising: with the development of technology and medical skills, we can now detect any potential factors that lead to fetus deaths more efficiently.

## Geographical Distribution
```{r}
#install.packages("rgdal",dependencies = TRUE)
library(choroplethrMaps)
library(choroplethr)
df_geo <- fetal_2 %>%
  dplyr::group_by(Standard.Residence.States) %>%
  dplyr::summarise(death = sum(Fetal.Deaths), death.p = sum(Percent.of.Total.Deaths)) %>%
  as.data.frame() %>%
  transmute(region = tolower(`Standard.Residence.States`), value = death.p)

state_choropleth(df_geo, title = "Contribution of Fetal Death to Total Death in US",
                 legend = "Percentage")
```
From the graph above, we can clearly see the contribution of different US states to the total number of US fetal deaths. We found that CA, TX, FL, GA, PA, NY and IL accounted for the largest share of total fetal deaths in the United States. The states with the highest number of fetal deaths are mostly along the coast. However, some northern states (ex. MT, ND, ME, VT...) have very low numbers of fetal deaths.

## Fetal Characteristics

### Infant Delivery Weights
```{r}
library(forcats)
fetal_2 %>% 
  na.omit() %>%
  dplyr::group_by(Infant.Delivery.Weight.14) %>%
  dplyr::summarise(add = sum(Percent.of.Total.Deaths)) %>%
  dplyr::arrange(add) %>%
  ggplot2::ggplot(aes(x = add, y = forcats::fct_reorder(Infant.Delivery.Weight.14,add))) +
  ggplot2::geom_point(color = "blue") +
  ggplot2::theme_linedraw() +
  ggplot2::ylab("Percent") +
  ggplot2::xlab("Infant Weight") +
  ggplot2::ggtitle("Percent of Total Deaths for Different Infant Weights")
```

According to the World Health Organization (WHO), the average weight at birth for a full-term baby boy is 3,300 grams (g). The average full-term baby girl weighs 3,200 g, at birth. The average baby born at 37 to 40 weeks weighs between 2,500 to 4,000 grams. Birth weights less than 2,500 grams will be considered low birth weights.

From the figure above, we can see that fetuses with a birth weight of 500-999 grams and fetuses with a birth weight of 350-499 grams died the most. The number of deaths of fetuses with birth weights in these two ranges was much higher than for any other birth weight. Not only that, we can also see that the top number of deaths of fetuses were under 2500 grams of birth weight. So we can conclude that underweight fetuses have a higher mortality rate. However, the data shows that the number of fetal deaths outside the average range of fetal weight is low.

### Plurality of Birth
```{r}
df_plur <- fetal_3 %>%
  group_by(Year, Plurality.or.Multiple.Birth) %>%
  summarise(add = sum(Fetal.Deaths)) %>%
  pivot_wider(names_from = Plurality.or.Multiple.Birth, values_from = add)
df_plur[is.na(df_plur)] <- 0
df_plur_new <- df_plur %>%
  mutate(`Quadruplet or higher` = `Quadruplet `+`Quintuplet or higher`+`Quadruplet or higher`) %>%
  select(c(-2,-3)) %>%
  pivot_longer(!Year,names_to = "Plurality", values_to = "Deaths") %>%
  group_by(Plurality, Year) %>%
  summarise(add = sum(Deaths))
df_plur_new$Plurality <- 
  factor(df_plur_new$Plurality, levels = c("Single", "Twin","Triplet","Quadruplet or higher"))

ggplot(data = df_plur_new, aes(x = Plurality, y = add)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.1) +
  facet_wrap( ~ Plurality, scales = "free") +
  ylab("Number of Fetal Deaths") +
  ggtitle("Distributions of Fetal Deaths in Multiple Birth")
```

In this part, we analyze whether the plurality of birth affects the number of fetal death. The above plot is a violin plot. The violin plot effectively shows not only median, range and variability, but also data distribution and its probability density. In the chart above, we can see that the single births have the widest distribution of fetal deaths and the highest number of fetal deaths.

### Infant Delivery Method
```{r}
# library(parcoords)
# library(GGally)
# library(ggalluvial)
# df_alluvial <- fetal_2 %>%
#   dplyr::group_by(Delivery.Method, Plurality.or.Multiple.Birth) %>%
#   dplyr::summarise(total = sum(Percent.of.Total.Deaths)) %>%
#   na.omit() 
# ggplot(df_alluvial, aes(axis1 = Delivery.Method, axis2 = Plurality.or.Multiple.Birth, 
#                         y = total)) +
#   geom_alluvium(color = "black" ) +
#   geom_stratum() +
#   geom_text(stat = "stratum", aes(label = paste(after_stat(stratum)))) +
#   scale_x_discrete(limits = c("Delivery Method","Multiple Birth")) +
#   ylab("Percentage to Total Death") +
#   ggtitle("Flow of change between delivery method and plurality of birth")

df_mosaic <- fetal_3%>%
  group_by(Year,Plurality.or.Multiple.Birth,Delivery.Method) %>%
  summarise(add = sum(Fetal.Deaths)) %>%
  pivot_wider(names_from = Plurality.or.Multiple.Birth, values_from = add) 
df_mosaic[,3:8][is.na(df_mosaic[,3:8])] <- 0

df_mosaic_new <- df_mosaic %>%
  mutate(`Quadruplet or higher` = `Quadruplet `+`Quintuplet or higher`+`Quadruplet or higher`) %>%
  select(c(-3,-4)) %>%
  pivot_longer(cols = -c(Year, Delivery.Method), names_to = "Pluarity", values_to = "Deaths") %>%
  na.omit()
df_mosaic_new$Pluarity <- factor(df_mosaic_new$Pluarity, 
                                 levels = c("Single","Twin","Triplet","Quadruplet or higher"))

tab_mosaic <- xtabs(Deaths~Delivery.Method + Pluarity,df_mosaic_new)
mosaicplot(tab_mosaic, direction = c("h","v"),
           main = "Delivery Method vs. Multiple Birth",las = 1,
           xlab = "Delivery Method", ylab = "Plurality", color = 
             c("#DEEBF7","#9ECAE1","#4292C6","#08306B"))
```
We plotted the relationship between different delivery methods, the number of fetuses per birth and the number of fetal deaths. Above is a Mosaic plot. From the graph above, we can clearly see that the number of fetal deaths caused by cesarean is much lower than that caused by vaginal delivery, no matter the fetus is single, twin, triplet or more. Moreover, singletons were the most likely to die, regardless of delivery method.

### Gestational Age
```{r}
df_bar <- fetal_2 %>%
  dplyr::group_by(LMP.Gestational.Age.10) %>%
  dplyr::summarise(add = sum(Percent.of.Total.Deaths)) %>%
  dplyr::arrange(desc(add)) %>%
  na.omit()

ggplot(data = df_bar,aes(x = reorder(LMP.Gestational.Age.10,add), y=add)) +
  geom_bar(stat = "identity",fill = "cornflowerblue") +
  ggtitle("LMP Gestational Age Distribution") +
  xlab("LMP Gestational Age") +
  ylab("Percent to Total Death") +
  coord_flip() +
  theme(panel.grid.major.x = element_blank())
```



## Maternal/Paternal Characteristics

```{r}
library(tidyverse)
library(ggplot2)
parentsAge <- read.delim("parentsAge.txt") %>% 
  dplyr::select(Age.of.Mother.9, Age.of.Father,Fetal.Deaths) %>%
  dplyr::filter(Age.of.Father != "" & Age.of.Father != "Unknown or Not Stated" & Age.of.Father != "Not Reported")
g_age <- ggplot(parentsAge, aes(x = Age.of.Father, y = Fetal.Deaths)) +
  geom_col(position = "dodge") +
  coord_flip() + 
  facet_wrap(~Age.of.Mother.9) + 
  ggtitle("Fetal Deaths Faceted by Maternal Age") + 
  xlab("Paternal Age") +
  ylab("# of Fetal Deaths")
print(g_age)
```

```{r}
#race <- fetal %>% dplyr::select(Mother.s.Bridged.Race, Year, Fetal.Deaths) %>%
#  filter(Mother.s.Bridged.Race != "Not Reported") %>%
#  group_by(Mother.s.Bridged.Race, Year) %>%
#  summarize(n = n())

#library(ggridges)
#g_race <- ggplot(race, aes(x = Year, y = Mother.s.Bridged.Race, fill = Mother.s.Bridged.Race)) +
#  geom_density_ridges(aes(n)) +
#  theme_ridges() + 
#  theme(legend.position = "none") +
#  ggtitle("")
```





### Clevland plot

```{r}
library(dplyr)
library(ggplot2)
Race_N_Marriage <- read.delim("RaceNMarriage.txt") %>% 
  rename('Maternal_Race' = Mother.s.Bridged.Race) %>%
  rename('Maternal Marital Status' = Mother.s.Marital.Status) %>% 
  na.omit() %>%
  dplyr::select(`Maternal_Race`, `Maternal Marital Status`, Fetal.Deaths) %>%
  filter(`Maternal Marital Status` == "Married" | `Maternal Marital Status` == "Unmarried") %>%
  group_by(Maternal_Race,'Maternal Marital Status')

ggplot(Race_N_Marriage, aes(x = Maternal_Race, y = Fetal.Deaths, color = `Maternal Marital Status`)) +
  geom_point(size = 3) +
  coord_flip() +
#  facet_wrap(~Maternal Marital Status)+
  ylab("# of Fetal Deaths") +
  xlab("Maternal Race")
```

