--- 
title: "Research on Fetal Deaths"
author: "Jiayi Yuan (jy3162) & Yayun Luo (yl4832)"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Introduction

```{r}
#install.packages("bookdown")
browseURL("docs/index.html")
```

The fetal mortality rate is the measure of poor reproductive health outcomes. It also indicates to some extent a country's medical and health level, economic level and basic implementation level. As the most advanced developed country in the world, the United States is famous for its advanced medical technology and strong economic strength. However, the United States ranks very high in fetal mortality. The fetal mortality rate in the United States is even three times that of some developed countries (such as Japan and Sweden), which shows that fetal mortality is a serious issue that needs to be taken seriously.


We believe that the death of a fetal is not only a sad tragedy for the family, but also has a very negative impact on an Aging American society. Therefore, in this project, we aim to explore the causes of fetal death. We will analyze the distribution of fetal mortality in the United States from the perspective of time and space. We will also examine whether specific characteristics of parents contribute to fetal mortality.

For more details of our project, please check out our [**GitHub repository**](https://github.com/MiraYuan/FetalDeath) or copy the following link https://github.com/MiraYuan/FetalDeath and paste it in your browser. 


<!--chapter:end:index.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data sources

All of our datasets were downloaded from the [**CDC Wonder API—Fetal Deaths, 2005-2019 Request**](https://wonder.cdc.gov/controller/datarequest/D133) (https://wonder.cdc.gov/controller/datarequest/D133). Since CDC Wonder only provides limited numbers of “group by” options, we needed to collect several datasets from the API. We first figuring out the variables we want to focus on and the relationships we want to study, and then downloaded informative datasets from the API.

## Dataset Documentation



<!--chapter:end:02-data.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data transformation

Since the data files downloaded from CDC Wonder are already in the tidy version, we do not need to do too much work on transforming the data. What we focused on was to figure out what values should be changed to missing values, for instance, values recorded as unknown or not stated.

```{r}
library(tidyr)
fetal <- read.delim("Fetal Deaths, 2005-2019.txt")
birth <- read.delim("Natality, 2007-2020.txt")
fetal <- fetal[,c(-1,-3,-5,-7,-9,-11)]
fetal <- na.omit(fetal)
birth <- birth[,c(-1,-3,-5,-7)]
birth <- na.omit(birth)
fetal$Age.of.Mother.9 <- as.factor(fetal$Age.of.Mother.9)

birth_s_g <- read.delim("Natality, 2007-2020 (2).txt")
birth_s_g <- birth_s_g[,c(-1,-3,-5,-7,-9)]
birth_s_g <- na.omit(birth_s_g)
fetal_2 <- read.delim("Fetal Deaths, 2005-2019 (1).txt") %>%
  dplyr::select(2,4,6,8,10,12,13)
fetal_2 <- fetal_2[1:3942,]
fetal_2 <- fetal_2 %>%
  dplyr::mutate(Infant.Delivery.Weight.14 = 
           ifelse(grepl("Not",Infant.Delivery.Weight.14, fixed = TRUE), NA_character_, Infant.Delivery.Weight.14)) %>%
  dplyr::mutate(LMP.Gestational.Age.10 = 
           ifelse(grepl("Not", LMP.Gestational.Age.10, fixed = TRUE), NA_character_, LMP.Gestational.Age.10)) %>%
  dplyr::mutate(Delivery.Method = 
           ifelse(grepl("Not", Delivery.Method, fixed = TRUE), NA_character_, Delivery.Method)) %>%
  dplyr::mutate(Plurality.or.Multiple.Birth = 
           ifelse(grepl("Not", Plurality.or.Multiple.Birth, fixed = TRUE),
                   NA_character_, Plurality.or.Multiple.Birth))

birth_state <- read.delim("Natality, 2007-2019 state.txt") %>%
  dplyr::select(2,4,6)
birth_state_new <- birth_state[1:663,] %>%
  dplyr::group_by(State) %>%
  dplyr::summarise(total_birth = sum(Births))

birth_weight <- read.delim("Natality, 2007-2019 weight.txt") %>%
  dplyr::select(2,4)
birth_weight <- birth_weight[1:13,]
```








<!--chapter:end:03-cleaning.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Missing values

## Fetus Death Dataset used for 

```{r}
library(redav)
plot_missing <- function(x, percent = TRUE) {	
  na_count_all <- data.frame(is.na(x)) %>%	
    dplyr::group_by_all() %>%	
    dplyr::count(name = "count", sort = TRUE) %>%	
    dplyr::ungroup() %>%	
    tibble::rownames_to_column("pattern")	
  
  na_count_all <- na_count_all %>% 
    dplyr::mutate(pattern = factor(.data$pattern, levels = nrow(na_count_all):1))
  
  # count the number of columns with missing values; will be used later to determine if there's a "none missing" pattern	
  na_count_all <- na_count_all %>% 	
    dplyr::rowwise() %>%	
    dplyr::mutate(num_missing_cols = sum(dplyr::c_across(where(is.logical))))	
  
  # data frame for missing patterns bar chart	
  na_count_by_pattern <- na_count_all[,c("pattern", "count", "num_missing_cols")]
  na_count_by_pattern$none_missing <- ifelse(na_count_by_pattern$num_missing_cols == 0, TRUE, FALSE)

  # data frame for missing by column bar chart	
  na_count_by_column <- data.frame(is.na(x)) %>%	
    colSums() %>% 	
    sort(decreasing = TRUE) %>% 	
    tibble::enframe(name = "var", value = "count")	
  
  # tidy and sort na_count_all by column counts	
  na_count_all_tidy <- na_count_all %>% 	
    tidyr::pivot_longer(where(is.logical), names_to = "variable") %>%	
    dplyr::mutate(variable = factor(.data$variable, levels = na_count_by_column$var))  %>% 	
    dplyr::mutate(none_missing = ifelse(.data$num_missing_cols == 0, TRUE, FALSE))	
  
  # main plot
  main_plot <- ggplot2::ggplot(na_count_all_tidy, 
                               ggplot2::aes(.data$variable,.data$pattern, 
                                            fill = factor(.data$value), 
                                            alpha = .data$none_missing)) +	
    ggplot2::geom_tile(color = "white") +	
    ggplot2::scale_fill_manual(values = c("grey70", "mediumpurple")) +	
    ggplot2::scale_alpha_manual(values = c(.7, 1)) +	
    ggplot2::ylab("missing pattern") +	
    ggplot2::guides(fill = "none", alpha = "none") +	
    ggplot2::theme_classic(12) +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size=7))
  
  # check for "none missing" pattern
  none_missing_pattern <- na_count_by_pattern %>%
    dplyr::filter(.data$none_missing) %>% dplyr::pull(.data$pattern)
  
  if (length(none_missing_pattern) > 0) {	
    main_plot <- main_plot +	
      ggplot2::annotate("text", x = (ncol(na_count_all)-2)/2,	
               y = nrow(na_count_all) + 1 - as.numeric(as.character(none_missing_pattern)),	
               label = "complete cases")	
  }	
  
  # margin plots
  
  denom <- ifelse(percent, nrow(x)/100, 1)
  
  missing_by_column_plot <- ggplot2::ggplot(na_count_by_column,
                                            ggplot2::aes(forcats::fct_inorder(.data$var),
                                                         .data$count/denom)) +	
    ggplot2::geom_col(fill = "cornflowerblue", alpha = .7) +
    ggplot2::scale_y_continuous(expand = c(0, 0), n.breaks = 3) +	
    ggplot2::xlab("") +
    ggplot2::ylab(ifelse(percent, "% rows \n missing:", "num rows \n missing:")) +	
    ggplot2::theme_linedraw(12) + 	
    ggplot2::theme(panel.grid.major.x = ggplot2::element_blank(),	
          panel.grid.minor.x = ggplot2::element_blank(),
          axis.text.x = ggplot2::element_text(angle = 90, size=7))	
  
  missing_by_pattern_plot <- 
    ggplot2::ggplot(na_count_by_pattern, 
                    ggplot2::aes(.data$pattern, .data$count/denom, 
                                 alpha = .data$none_missing)) +
    ggplot2::geom_col(fill = "cornflowerblue") +
    ggplot2::coord_flip() +
    ggplot2::scale_y_continuous(expand = c(0, 0), n.breaks = 3) +
    ggplot2::scale_alpha_manual(values = c(.7, 1)) +
    ggplot2::xlab("") +
    ggplot2::ylab(ifelse(percent, "% rows", "row count")) +
    ggplot2::guides(alpha = "none") +
    ggplot2::theme_linedraw(12) +
    ggplot2::theme(panel.grid.major.y = ggplot2::element_blank(), 
          panel.grid.minor.y = ggplot2::element_blank())
  
  if (percent) {	
    missing_by_column_plot <- missing_by_column_plot +
      ggplot2::scale_y_continuous(expand = c(0, 0), n.breaks = 5,
                         limits = c(0, 100))	
    missing_by_pattern_plot <- missing_by_pattern_plot +
      ggplot2::scale_y_continuous(expand = c(0, 0), n.breaks = 5,
                         limits = c(0, 100))	
  }
  
  #par(mfrow = c(2, 2))
  list(main_plot, missing_by_column_plot, missing_by_pattern_plot)
  missing_by_column_plot + patchwork::plot_spacer() +
    main_plot + missing_by_pattern_plot +
    patchwork::plot_layout(widths = c(4, 1), heights = c(1, 4))
 }

#utils::globalVariables("where")
```

```{r comment=NA}
plot_missing(fetal_2)
# f1 <- sum(is.na(fetal))
# b1 <- sum(is.na(birth))
# b2 <- sum(is.na(birth_s_g))
# df <- data.frame(Dataset = c("Fetal Death", "Natality", "Natality with gender and state"),
#            "Missing Values" = c(f1, b1, b2))
# knitr::kable(df)
```


<!--chapter:end:04-missing.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Results

## Fetal Deaths Over Years
```{r}
library(ggplot2)
library(dplyr)
library(plotly)

### Total
g1 <- fetal %>%
  dplyr::group_by(Year) %>%
  dplyr::summarise(add = sum(Fetal.Deaths), add_p = sum(Percent.of.Total.Deaths)) %>%
  ggplot(aes(x = Year, y = add_p)) +
  geom_line() +
  geom_point(alpha = 0.5) +
  geom_point(aes(x = 2007, y = 7.22), color = "red") +
  geom_text(aes(label = add_p), nudge_x = 0.4, nudge_y = -0.06, size = 3) +
  labs(x = "Year", y = "Total Fetal Death Rate",
       title = "Fetal Death Rate Over Years") +
  scale_x_continuous(breaks = seq(2005, 2019, 1))
g1


g2 <- fetal %>%
  dplyr::group_by(Sex, Year) %>%
  dplyr::summarise(add_g = sum(Fetal.Deaths), add_g_p = sum(Percent.of.Total.Deaths)) %>%
  ggplot(aes(x = Year, y = add_g_p, col = Sex)) +
  geom_line() +
  geom_point(alpha = 0.5) +
  labs(x = "Year", y = "Fetal Death Rate",
       title = "Fetal Death Rate by Different Genders Over Years") +
  scale_x_continuous(breaks = seq(2005, 2019, 1))

start = 2007
end = 2008
g2 + annotate("rect", xmin = start, xmax = end, ymin = -Inf, ymax = Inf, 
              fill = "green", alpha = .2) +
  annotate("text", x = 2008.1, y = 3, label = "2007-2008", 
           color = "black", hjust = 0) +
  theme_classic()

# g3 <- birth_s_g %>%
#   group_by(Gender, Year) %>%
#   summarise(sum = sum(Births)) %>%
#   ggplot(aes(x = Year, y = sum, col = Gender)) +
#   geom_line() +
#   geom_point() +
#   labs(x = "Year", y = "Number of Births", 
#        title = "Number of Births by Different Genders Over Years") +
#   theme_classic() +
#   scale_x_continuous(breaks = seq(2007, 2019, 1))
# g3

# fetal %>%
#   group_by(Year) %>%
#   summarise(add = sum(Fetal.Deaths), add_p = sum(Percent.of.Total.Deaths)) %>%
#   plot_ly(x = ~Year, y = ~add, type = "scatter", mode = "lines+markers",
#           hoverinfo = "text", text = ~add) %>%
#   layout(title = "Total Fetal Deaths Over Years", 
#          xaxis = list(title = "Year", breaks = seq(2005, 2019, 1)),
#          yaxis = list(title = "Number of Total Death"))
```

First of all, let us take a look at how the fetal death rates changed between 2005 and 2019. We can clearly see a decreasing trend since 2007. Since we don’t have data before 2007, even though our graphs show there might be an increasing trend before 2007, we cannot simply make a conclusion. After further checking the movement by splitting the data into different genders, we found out that from 2007 to 2008, the female fetal death rate had a sharper decrease than male data did. We should notice that our graph shows higher death rates for male fetuses. This is not because male fetuses have a higher possibility of death but because male fetuses have higher birth rates in general. This decreasing trend is a good indicator, showing an improvement in national health. It is also unsurprising: with the development of technology and medical skills, we can now detect any potential factors that lead to fetus deaths more efficiently.

## Geographical Distribution
```{r}
library(choroplethrMaps)
library(choroplethr)
df_geo <- fetal_2 %>%
  dplyr::group_by(Standard.Residence.States) %>%
  dplyr::summarise(death = sum(Fetal.Deaths), death.p = sum(Percent.of.Total.Deaths)) %>%
  as.data.frame() %>%
  transmute(region = tolower(`Standard.Residence.States`), value = death.p)

state_choropleth(df_geo, title = "Contribution of Fetal Death to Total Death in US",
                 legend = "Percentage")
```
install.packages("rgdal",dependencies = TRUE)

## Remarkable Fetal Characteristics

### Does Infant Weights matters?
```{r}
library(forcats)
fetal_2 %>% 
  na.omit() %>%
  dplyr::group_by(Infant.Delivery.Weight.14) %>%
  dplyr::summarise(add = sum(Percent.of.Total.Deaths)) %>%
  dplyr::arrange(add) %>%
  ggplot2::ggplot(aes(x = add, y = forcats::fct_reorder(Infant.Delivery.Weight.14,add))) +
  ggplot2::geom_point(color = "blue") +
  ggplot2::theme_linedraw() +
  ggplot2::ylab("Percent") +
  ggplot2::xlab("Infant Weight") +
  ggplot2::ggtitle("Percent of Total Deaths for Different Infant Weights")
```

### Relationships between delivery method and plurality of birth
```{r}
library(parcoords)
library(GGally)
library(ggalluvial)
df_alluvial <- fetal_2 %>%
  dplyr::group_by(Delivery.Method, Plurality.or.Multiple.Birth) %>%
  dplyr::summarise(total = sum(Percent.of.Total.Deaths)) %>%
  na.omit() 
ggplot(df_alluvial, aes(axis1 = Delivery.Method, axis2 = Plurality.or.Multiple.Birth, 
                        y = total)) +
  geom_alluvium(color = "black" ) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = paste(after_stat(stratum)))) +
  scale_x_discrete(limits = c("Delivery Method","Multiple Birth")) +
  ylab("Percentage to Total Death") +
  ggtitle("Flow of change between delivery method and plurality of birth")
```

```{r}
df_mosaic <- fetal_2 %>%
  dplyr::group_by(Delivery.Method,Plurality.or.Multiple.Birth) %>%
  dplyr::summarise(add = sum(Percent.of.Total.Deaths)) %>%
  na.omit()

df_mosaic$Plurality.or.Multiple.Birth <- factor(df_mosaic$Plurality.or.Multiple.Birth,
                                             levels = c("Single","Twin","Triplet"))
  

tab_mosaic <- xtabs(add~.,df_mosaic)

vcd::mosaic(Delivery.Method~Plurality.or.Multiple.Birth, tab_mosaic,
            direction = c("v","h"), set_labels=list(Plurality.or.Multiple.Birth 
                                                    = c("1", "2",">3")),
            gp_labels = grid::gpar(fontsize = 8), highlighting_fill = c("#D1E5F0","#2166AC"),
            main = "Delivery Method vs. Multiple Birth Number")
  
```

### LMP weeks
```{r}
df_bar <- fetal_2 %>%
  dplyr::group_by(LMP.Gestational.Age.10) %>%
  dplyr::summarise(add = sum(Percent.of.Total.Deaths)) %>%
  dplyr::arrange(desc(add)) %>%
  na.omit()

ggplot(data = df_bar,aes(x = reorder(LMP.Gestational.Age.10,add), y=add)) +
  geom_bar(stat = "identity",fill = "cornflowerblue") +
  ggtitle("LMP Gestational Age Distribution") +
  xlab("LMP Gestational Age") +
  ylab("Percent to Total Death") +
  coord_flip() +
  theme(panel.grid.major.x = element_blank())
```



## Maternal/Paternal Characteristics

```{r}
library(tidyverse)
library(ggplot2)
parentsAge <- read.delim("parentsAge.txt") %>% 
  dplyr::select(Age.of.Mother.9, Age.of.Father,Fetal.Deaths) %>%
  dplyr::filter(Age.of.Father != "" & Age.of.Father != "Unknown or Not Stated" & Age.of.Father != "Not Reported")
g_age <- ggplot(parentsAge, aes(x = Age.of.Father, y = Fetal.Deaths)) +
  geom_col(position = "dodge") +
  coord_flip() + 
  facet_wrap(~Age.of.Mother.9) + 
  ggtitle("Fetal Deaths Faceted by Maternal Age") + 
  xlab("Paternal Age") +
  ylab("# of Fetal Deaths")
print(g_age)
```

```{r}
race <- fetal %>% dplyr::select(Mother.s.Bridged.Race, Year, Fetal.Deaths) %>%
  filter(Mother.s.Bridged.Race != "Not Reported") %>%
  group_by(Mother.s.Bridged.Race, Year) %>%
  summarize(n = n())

library(ggridges)
g_race <- ggplot(race, aes(x = Year, y = Mother.s.Bridged.Race, fill = Mother.s.Bridged.Race)) +
  geom_density_ridges(aes(n)) +
  theme_ridges() + 
  theme(legend.position = "none") +
  ggtitle("")
print(g_race)
```





### Clevland plot

```{r}
library(dplyr)
library(ggplot2)
Race_N_Marriage <- read.delim("RaceNMarriage.txt") %>% 
  rename('Maternal_Race' = Mother.s.Bridged.Race) %>%
  rename('Maternal Marital Status' = Mother.s.Marital.Status) %>% 
  na.omit() %>%
  dplyr::select(`Maternal_Race`, `Maternal Marital Status`, Fetal.Deaths) %>%
  filter(`Maternal Marital Status` == "Married" | `Maternal Marital Status` == "Unmarried") %>%
  group_by(Maternal_Race,'Maternal Marital Status')

ggplot(Race_N_Marriage, aes(x = Maternal_Race, y = Fetal.Deaths, color = `Maternal Marital Status`)) +
  geom_point(size = 3) +
  coord_flip() +
#  facet_wrap(~Maternal Marital Status)+
  ylab("# of Fetal Deaths") +
  xlab("Maternal Race")
```


<!--chapter:end:05-results.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Interactive component
Understand fetal death by region

## The Total Amount of Fetal Death from 2005 to 2019

```{r}
library(dygraphs)
library(xts)          
library(tidyverse)
library(lubridate)

#data
fetal_by_year <- fetal %>% group_by(Year) %>%
  summarize(`Total Fetal Deaths` = sum(Fetal.Deaths))

#plot
g_by_year <- dygraph(fetal_by_year) %>%
  dyOptions(labelsUTC = TRUE, fillGraph=TRUE, fillAlpha=0.1, drawGrid = FALSE, colors="#D8AE5A") %>%
  dyRangeSelector() %>%
  dyCrosshair(direction = "vertical") %>%
  dyHighlight(highlightCircleSize = 5, highlightSeriesBackgroundAlpha = 0.2, hideOnMouseOut = FALSE)  %>%
  dyRoller(rollPeriod = 1) 
g_by_year
```



## Changes in the Number of Fetal Deaths Among Different Maternal Ages in Different Years
```{r}
library(streamgraph)
library(tidyverse)

data <- fetal %>% group_by(Age.of.Mother.9, Year) %>% 
  summarise(Deaths = sum(Fetal.Deaths))
 
# Stream graph with a legend
g_age_year <- streamgraph(data, key="Age.of.Mother.9", value="Deaths", date="Year") %>%
  sg_legend(show=TRUE, label="names: ")
g_age_year
```


<!--chapter:end:06-interactive.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Conclusion


<!--chapter:end:07-conclusion.Rmd-->

